Galois Slicing as Differentiation

### 2025-07-11 main/approx-as-tangents.tex
Brackets for consistency with notation elsewhere.

We end this section with a conjecture. Although we have argued above
that there is an analogy between stable functions and smooth
functions, we have not stated any mathematical theorems substantiating
this. \emph{Tangent Categories} \cite{cockett14,cockett18} are a
categorical axiomatisation of the properties of manifolds and smooth
functions in terms of the presence of tangent bundles $TX$ for every
object $X$, forward derivatives and additivity of tangents. They
generalise Cartesian Differential Categories \cite{cdcs}, which are an
axiomatisation of Euclidean spaces and smooth functions.

\begin{conjecture}
  \label{con:tangent-stable-fns}
  (1) Bounded meet posets and conditionally multiplicative functions
  form a Tangent category where the tangent bundle
  $TX = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
  by meets. (2) L-posets and stable functions form a \emph{reverse
    Tangent category} \cite{reverse-tangents}.
\end{conjecture}

>>>>

We end this section with a conjecture. Although we have argued above
that there is an analogy between stable functions and smooth
functions, we have not stated any mathematical theorems substantiating
this. \emph{Tangent Categories} \cite{cockett14,cockett18} are a
categorical axiomatisation of the properties of manifolds and smooth
functions in terms of the presence of tangent bundles $T(X)$ for every
object $X$, forward derivatives and additivity of tangents. They
generalise Cartesian Differential Categories \cite{cdcs}, which are an
axiomatisation of Euclidean spaces and smooth functions.

\begin{conjecture}
  \label{con:tangent-stable-fns}
  (1) Bounded meet posets and conditionally multiplicative functions
  form a Tangent category where the tangent bundle
  $T(X) = \{(x,x') \mid x' \leq x\}$ and addition of tangents is given
  by meets. (2) L-posets and stable functions form a \emph{reverse
    Tangent category} \cite{reverse-tangents}.
\end{conjecture}

### 2025-07-11 main/introduction.tex
Brackets for consistency with notation elsewhere.

Existing approaches to program slicing are often tied to particular programming languages or implementations.
In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
called \GPS, where the forward and backward slicing procedures for a particular execution form a Galois
connection. Our main insight is that this approach to slicing can be seen as a generalised form of
differentiable programming, with approximations (slices) of elements playing the role of tangents.

>>>>

Existing approaches to program slicing are often tied to particular programming languages or implementations.
In this paper we develop a general categorical approach to program slicing, focusing on a particular technique
called \GPS, where the set of slices of a given value form a lattice of approximations and the forward and
backward slicing procedures for a particular execution induce Galois connections between these lattices. Our
main insight is that this approach to slicing can be seen as a generalised form of differentiable programming,
with approximations (slices) of elements playing the role of tangents. Our categorical semantics should
provide a foundation for providing
